---
title: "Regression analysis"
author: "nidie"
date: "2022-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(GGally)
library(tidyverse)
library(MASS)
library(car)
library(stringr)
```


```{r}
rent_listings <- read.csv("./data/rent_listings_data_cleaned.csv")
rent_listings$Property.Name. <- toupper(rent_listings$Property.Name.)
rent_listings <- unique(rent_listings[c("Property.Name.","Property.Type.","Asking.","Area.","travel_time_changi","travel_time_orchard","travel_time_raffles","District.","Bedrooms.","Bathrooms.","Asking.","Area.Builtup","is_HDB","Built.Year.","PSF.")])
data_rent <- read.csv("./Data/rent_transactions_private_properties cleaned.csv")
data_loc <- read.csv("./Data/districts_metadata.csv")
cte <- left_join(data_rent, data_loc, by = c("Postal.District" = "id"))
cte <- cte %>% mutate(lower = str_extract(cte$Floor.Area..sq.ft., "^\\d+(?=\\s)")) %>% mutate(upper = str_extract(cte$Floor.Area..sq.ft., "(?<=\\s)\\d+"))
agg_data <- left_join(cte, rent_listings, by = c("Building.Project.Name" = "Property.Name.")) %>% filter(Area.Builtup >= lower & Area.Builtup <= upper)
```


# Overall analysis
## Investigating relationship between area of the property and its prices across all types
```{r}
# Have a glimpse of the general pattern
ggplot(rent_listings, aes(x = Area.Builtup, y = Asking., col = Property.Type.)) + geom_point(size = 0.8)

## Detached, Semi-Detached, and Terraces are considerably larger in size and price

# Have a better understanding of common properties
listing_cleaned <- rent_listings %>% filter(rent_listings$Asking. < 20000 & rent_listings$Area. < 2500)
ggplot(data = listing_cleaned, aes(x = Area.Builtup, y = Asking., col = Property.Type.)) + geom_point(size = 0.8)

# HDBs - Smaller in size and better in price
# Apartment, Condominium, Terrace and Semi-detached needs deserves more analysis
```

## Distance from the central
```{r}
ggplot(agg_data, aes(x=travel_time_minutes_to_central, y=Monthly.Gross.Rent...)) + geom_point(size = 0.8) + geom_smooth(method = "lm")
```

## Multivariate regression
### Spliting training and validating use data
```{r}
agg_data_cleaned <- agg_data %>% filter(!is.na(is_HDB))
agg_data_cleaned$is_HDB <- case_when(agg_data_cleaned$is_HDB == "No" ~ 0, TRUE ~ 1)
total <- c(1:nrow(agg_data_cleaned))
tindex <- sample(total, as.integer(length(total)*0.75))
training <- agg_data_cleaned[tindex,]
vindex <- total[is.na(pmatch(total, tindex))]
validator <- agg_data_cleaned[vindex,]
```

### Deriving regression formula
```{r}
lmRent <- lm(PSF.~travel_time_minutes_to_central + travel_time_orchard + travel_time_raffles + Bathrooms. + Bedrooms. + Built.Year. + Area.Builtup + is_HDB, data = agg_data_cleaned)
summary(lmRent)
```

### Refining regressors with stepAIC
```{r}
stepRent <- stepAIC(lmRent, direction = "both")
summary(stepRent)
vif(stepRent)
```

### Predicting and validating regression result
```{r}
predictRent <- predict(stepRent, validator)
validator$prediction <- predictRent
ggplot(validator, aes(x=PSF., y=prediction)) + geom_point() + xlab("Actual Monthly Rent") + ylab("Predicted Rent") + geom_smooth(method = "lm")
```


## Apartment and Condominium comparison
```{r}
listing_C_A <- listing_cleaned %>% filter(listing_cleaned$Property.Type. %in% c("Condominium", "Apartment") & listing_cleaned$Area.Builtup >= 300)

ggplot(listing_C_A, aes(x = Area.Builtup, y = Asking., col = Property.Type.)) + geom_point(size = 0.8) + geom_smooth(method = "lm")


# Apartment and Condominium have almost identical pattern
```

## Detached, Semi-detached, and Terrace comparison
```{r}
# excluding outliers for the sake of LSA #3
listing_D_S_T <- rent_listings %>% filter(rent_listings$Property.Type. %in% c("Detached", "Semi-Detached", "Terrace") & rent_listings$Asking. <= 100000)

ggplot(listing_D_S_T, aes(x = Area.Builtup, y = Asking., col = Property.Type.)) + geom_point(size = 0.8) + geom_smooth(method = "lm")

# Almost Equally influenced by Property Area holding other variables constant
```

## Non-landed properties breakdown
### Non-landed properties near Changi
```{r}
# To better investigate other variables' influences we need multi-regressor regression
aptm <- agg_data_cleaned %>% filter(Type == "Executive HDB" | Type == "Non-landed Properties") %>% dplyr::select(District., PSF., Built.Year., Area.Builtup, travel_time_changi, travel_time_orchard, travel_time_raffles, Bathrooms., Bedrooms.)

aptm$District. <- as.integer(str_extract(aptm$District., "\\d+"))

# Near Changi
aptm_changi <- aptm %>% filter(District. >= 14 & District. <= 19) %>% dplyr::select(PSF., Built.Year., Area.Builtup, travel_time_changi, Bathrooms., Bedrooms.)

mod_changi <- lm(PSF. ~ Built.Year. + Area.Builtup + travel_time_changi + Bathrooms. + Bedrooms., method = "lm", data = aptm_changi)
summary(mod_changi)

stepChangi <- stepAIC(mod_changi, direction = "both")
summary(stepChangi)
vif(stepChangi)

## With an adjusted R squared of 0.9716, we are confident that the estimate does fit the dataset
## We can clearly see a very strong beta1 estimate in built year and area, which denotes that for apartments that are 1 year newer, it is 5.06 more expensive. For each additional sqft there is an increase in price of 4.64. We can also observe a negative correlation between travel time to Changi Airport and the rent price, which makes sense by common understanding.
```

### Non-landed properties near Orchards and Raffles
```{r}
# Near Orchard
aptm_orchard <- aptm %>% filter(District. >= 1 & District. <= 13) %>% dplyr::select(PSF., Built.Year., Area.Builtup, travel_time_orchard, Bedrooms., Bathrooms.)

## Variables show no clear correlations except the dependent variable Asking.. We are safe to continue our regression

mod_orchard <- lm(PSF. ~ Built.Year. + Area.Builtup + travel_time_orchard + Bedrooms. + Bathrooms., method = "lm", data = aptm_orchard)
summary(mod_orchard)

stepOrchard <- stepAIC(mod_orchard, direction = "both")
summary(stepOrchard)
vif(stepOrchard)

## We see an even stronger relationship between Area.Builtup and the rent price in regions near Orchards.
```

### Apartments near Raffles
```{r}
# Near Orchard
aptm_raffles <- aptm %>% filter(District. >= 20) %>% dplyr::select(PSF., Built.Year., Area.Builtup, travel_time_raffles, Bathrooms., Bedrooms.)


## Variables show no clear correlations except the dependent variable Asking.. We are safe to continue our regression

mod_raffles <- lm(PSF. ~ Built.Year. + Area.Builtup + travel_time_raffles + Bathrooms. + Bedrooms., method = "lm", data = aptm_raffles)
summary(mod_raffles)

stepRaffles <- stepAIC(mod_raffles, direction = "both")
summary(stepRaffles)
vif(stepRaffles)

## Raffles are not that heavily influenced by builtup area and built year compared to other regions near Orchards and Changi
```